---
title: "Phone_Usage"
author: "Jesse Stanley"
output: html_document
date: "2023-09-14"
---

## Setup & Cleaning
```{r}
#import packages
library(readxl)
library(dplyr)
library(chron)
library(ggplot2)
library(lubridate)
library(data.table)
library(ggthemes)
library(knitr)
library(kableExtra)
library(magick)

#read file
data = read_excel("CleanedData.xlsx")

#convert times to numeric
data$t_total = period_to_seconds(hms(format(data$t_total, 
                                            format = '%H:%M:%S')))/(60*60)
data$t_app = period_to_seconds(hms(format(data$t_app, 
                                          format = '%H:%M:%S')))/(60*60)
data$t_com_app = period_to_seconds(hms(format(data$t_com_app, 
                                              format = '%H:%M:%S')))/(60*60)
```

## View Raw Data
```{r}
View(data)
```

## Overview
```{r}
#plot total time
ggplot(data, aes(x=t_total)) +
  geom_bar(colour = "black", fill = "grey") +
  scale_x_binned(n.breaks = 20, show.limits = TRUE) +
  labs(title = "Total Time")

#plot most used apps
ggplot(data, aes(x=app)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.25), 
        axis.line.y = element_line()) +
    labs(title = "Most Used App")

#plot most used apps time
ggplot(data, aes(x=t_app)) +
  geom_bar(colour = "black", fill = "grey") +
  scale_x_binned(n.breaks = 20, show.limits = TRUE) +
  labs(title = "Most Used App Time")

#plot most used communication apps
ggplot(data, aes(x=com_app)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.25), 
        axis.line.y = element_line()) +
  labs(title = "Most Used Communication App")

#plot most used communication apps time
ggplot(data, aes(x=t_com_app)) +
  geom_bar(colour = "black", fill = "grey") +
  scale_x_binned(n.breaks = 20, show.limits = TRUE) +
  labs(title = "Most Used Communication App Time")
```

## Group Data
```{r}
#separate into Science & Engineering and Other groups
fgroup = data
fgroup[fgroup$faculty != 'Faculty of Science & Engineering',]['faculty'] = 
  'Other faculties'

#sample group sizes
scieng = count(fgroup[fgroup$faculty == "Faculty of Science & Engineering",])
oth = count(fgroup[fgroup$faculty == "Other faculties",])
cat("Sample Sizes\n\nScience and eng: ",as.character(scieng),"\nOther: ",
    as.character(oth),"\n")
```

## Hitograms of Total Time 
```{r}
#stacked together
ggplot(fgroup, aes(x=t_total, fill=faculty)) +
  geom_bar(colour = "black") +
  scale_x_binned(n.breaks = 20, show.limits = TRUE) +
  scale_fill_manual(values = c("#F8766D","#619CFF")) +
  labs(caption = bquote(bold("Figure 2.\n")~"Hours of phone usage on weekdays of Curtin University students from the Faculty of Science & Engineering\nand other faculties showing approximate normal distribution.")) +
  xlab("Time (hours)") +
  ylab("Number of Students") +
  theme(plot.caption = element_text(vjust = -2), plot.caption.position = "panel", legend.position = "none") +
  facet_grid(faculty ~ .)
  
#Faculty of Science & Engineering
ggplot(fgroup[fgroup$faculty == "Faculty of Science & Engineering",], aes(x=t_total)) +
  geom_bar(colour = "black", fill = "#F8766D") +
  scale_x_binned(n.breaks = 20, show.limits = TRUE, limits = c(0,20)) +
  labs(title = "Total Phone Usage for Science & Engineering Faculty", 
       caption = "Hours of phone usage on weekday for Science & Engineering students") +
  xlab("Hours") +
  ylab("Number of Students")

#Other Faculties
ggplot(fgroup[fgroup$faculty == "Other faculties",], aes(x=t_total)) +
  geom_bar(colour = "black", fill = "#619CFF") +
  scale_x_binned(n.breaks = 20, show.limits = TRUE, limits = c(0,20))
```

## Testing Total Time for Assosciation
```{r}
#F-test
var.test(t_total ~ faculty, data = fgroup)
```

p-value > 0.05, therefore the null hypothesis that the variances\
are the same is a safe assumption

```{r}
#t-test
t.test(t_total ~ faculty, data = fgroup, detailed = TRUE, var.equal = TRUE)
```

p-value < 0.05, therefore the alternate hypothesis accepted that the average\
number of hours spent on a phone on a weekday is different for Science and\
engineering students compared to other faculties

## Boxplots of Total Time
```{r}
ggplot(fgroup, aes(x=faculty, y=t_total, fill=faculty)) +
  stat_boxplot(geom = "errorbar", width = .25) +
  geom_boxplot(show.legend = FALSE) +
  scale_fill_manual(values = c("#F8766D","#619CFF")) +
  labs(caption = bquote(bold("Figure 1.\n")~"Mean, median, and IQR of weekday phone usage of Curtin University students\nfrom the Faculty of Science & Engineering and other faculties.")) +
  theme_hc() +
  theme(plot.caption = element_text(vjust = -2.5)) +
  stat_summary(fun = "mean", shape = 1, show.legend = FALSE) +
  xlab("Faculty") +
  ylab("Time (hours)")
```

## Histograms of Most Used App Time
```{r}
#stacked together
ggplot(fgroup, aes(x=t_app, fill=faculty)) +
  geom_bar(colour = "black", position = "stack") +
  scale_x_binned(n.breaks = 10, show.limits = TRUE, limits = c(0,10)) +
  scale_fill_manual(values = c("#F8766D","#619CFF"))

#Faculty of Science & Engineering
ggplot(fgroup[fgroup$faculty == "Faculty of Science & Engineering",], aes(x=t_app)) +
  geom_bar(colour = "black", fill = "#F8766D") +
  scale_x_binned(n.breaks = 10, show.limits = TRUE, limits = c(0,10))

#Other Faculties
ggplot(fgroup[fgroup$faculty == "Other faculties",], aes(x=t_app)) +
  geom_bar(colour = "black", fill = "#619CFF") +
  scale_x_binned(n.breaks = 10, show.limits = TRUE, limits = c(0,10))
```

```{r}
#Barplot of most used apps
ggplot(fgroup, aes(x=app, fill=faculty)) +
  geom_bar(position = position_dodge2(preserve = "single", padding = 0.05)) +
  scale_fill_manual(values = c("#F8766D","#619CFF")) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.25))
```

## Testing Most Used App Time for Association
```{r}
#F-test
var.test(t_app ~ faculty, data = fgroup)
```

p-value > 0.05, therefore the null hypothesis that the variances\
are the same is a safe assumption

```{r}
#t-test
t.test(t_app ~ faculty, data = fgroup, detailed = TRUE, var.equal = TRUE)
```

p-value > 0.05, therefore the null hypothesis is accepted that the\
average number of hours spent on the most used app on a weekday is the same for\
Science and engineering students compared to other faculties

## Boxplots of Most Used App Time
```{r}
ggplot(fgroup, aes(x=faculty, y=t_app)) +
  geom_boxplot()
```

## Histograms of Most Used Communication App Time
```{r}
#stacked together
ggplot(fgroup, aes(x=t_com_app, fill=faculty)) +
  geom_bar(colour = "black", position = "stack") +
  scale_x_binned(n.breaks = 10, show.limits = TRUE, limits = c(0,10)) +
  scale_fill_manual(values = c("#F8766D","#619CFF"))

#Faculty of Science & Engineering
ggplot(fgroup[fgroup$faculty == "Faculty of Science & Engineering",], aes(x=t_com_app)) +
  geom_bar(colour = "black", fill = "#F8766D") +
  scale_x_binned(n.breaks = 10, show.limits = TRUE, limits = c(0,10))

#Other Faculties
ggplot(fgroup[fgroup$faculty == "Other faculties",], aes(x=t_com_app)) +
  geom_bar(colour = "black", fill = "#619CFF") +
  scale_x_binned(n.breaks = 10, show.limits = TRUE, limits = c(0,10))
```
```{r}
#Barplot of most used communication app
ggplot(fgroup, aes(x=com_app, fill=faculty)) +
  geom_bar(position = position_dodge2(preserve = "single", padding = 0.05)) +
  scale_fill_manual(values = c("#F8766D","#619CFF")) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.25))
```

## Testing Most Used Communication App Time for Association
```{r}
#F-test
var.test(t_com_app ~ faculty, data = fgroup)
```

p-value > 0.05, therefore the null hypothesis that the variances\
are the same is a safe assumption

```{r}
#t-test
t.test(t_com_app ~ faculty, data = fgroup, detailed = TRUE, var.equal = TRUE)
```

p-value > 0.05, therefore the null hypothesis is accepted that the\
average number of hours spent on the most used communication app on a weekday\
is the same for Science and engineering students compared to other faculties

## Boxplots for Most Used Communication App Time
```{r}
#boxplots for most used com app
ggplot(fgroup, aes(x=faculty, y=t_com_app)) +
  geom_boxplot()
```

## Statistics Table
```{r}
eng_t = fgroup$t_total[fgroup$faculty == "Faculty of Science & Engineering"]
oth_t = fgroup$t_total[fgroup$faculty != "Faculty of Science & Engineering"]
tab = data.frame("N" = c(length(eng_t),length(oth_t)), 
                 "mean" = c(mean(eng_t), mean(oth_t)), 
                 "sd" = c(sd(eng_t),sd(oth_t)), 
                 row.names = c("Faculty of Science & Engineering", 
                               "Other Faculties"))

kable(tab, caption = "Table 1. sample size, mean and standard deviation of Curtin University students from seperate groups", 
      digits = 2) %>%
  kable_classic(full_width = F) %>%
  as_image()
```

## Other Statistics
```{r}
#count of students whos most used app is the same as most used communication app
secount = count(fgroup[fgroup$faculty == 'Faculty of Science & Engineering' & fgroup$com_app == "Same as most used app",])
ocount = count(fgroup[fgroup$faculty == 'Other faculties' & fgroup$com_app == "Same as most used app",])
cat('"Same as most used app count"\n\nScience and eng: ',as.character(secount),"\nOther: ",as.character(ocount),"\n")

#summary
vtable::sumtable(fgroup)
```

